Spike spectra for recurrences

### 2022-11-03 Manuscript/inter_spike_manuscript_mdpi.tex

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In reccurance analysis, the $\tau$-recurrence rate encodes the periods of the cycles of the underlying high dimensional time series. It thus plays a similar role to the autocorrelation for scalar time series in encoding temporal correlations. Unlike the autocorrelation however, its Fourier decomposition does not have a natural interpretation. Thus there is no direct analogue to the power spectrum in recurrence analysis. In this paper we introduce a novel way to decompose the $\tau$-recurrence rate using an overcomplete basis of Dirac combs together with a sparsity regularization. We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis in the sense that it reveals the dominant periodicities of the underlying time series. We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying system in a wide variety of examples and is robust to measurement noise.}
%
%A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
%clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
%We invent this instructive approach with the 
%idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
%along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
%spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}

@@

\citet{Zbilut2008} pointed out that $\tau$-RR could be interpreted as analogous to the auto-correlation function $C(\tau)$ for a scalar time series. However, as there is no equivalent of the Wiener-Khinchin theorem its Fourier Transform (FT) only provides a ``generalized'' kind of power spectral density. There are clear advantages to such a recurrence-derived 
power spectral density, the (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D,E): There are no assumptions for stationarity or sampling, when constructing a RP, and the approch directly applies to higher-dimensional timeseries for which the auto-correlation is a matrix.\\

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Fig.~\ref{fig_tau_rr_dirac_comb}A, D), so 
in principle we would strive for a single peak in the frequency domain of this signal at a frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many frequencies excited (Fig.~\ref{fig_tau_rr_dirac_comb}B, E). This is because the Fourier components add constructively 
for every frequency $1/T_\text{is}$ and therefore $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$.

@@

In this article we propose a new way of decomposing a spike-train-like signal into periodic components as 
an alternative to the RP-based method suggested in \cite{Zbilut2008} or Fourier-based spike-train power 
spectra \cite{Dummer2014}. We instead investigate decomposing the $\tau$-RR in an overcomplete basis of spike trains using a sparsity regularization. Section \ref{sec_tau_rr_method} introduces the approach in detail. Our novel \textit{inter-spike spectrum} does not show resonance 
behavior of the signal's inherent inter-spike frequencies (Fig.~\ref{fig_tau_rr_dirac_comb}C, F). Note that this approach can be used to decompose arbitrary signals, and is not specifc to $\tau$-RR. The more spiky the signal is, the more useful our new decomposition can potentially be. In Section \ref{sec_tau_rr_application} we demonstrate the usage of inter-spike spectrum when transforming the $\tau$-RR of a system under study. In this case, the inter-spike spectrum 
can unravel characteristic time scales of high dimensional systems, which is not possible when using a FT. Finally, in Section \ref{sec_tau_rr_conclusion} our results are summarized.

@@

The signal, which we would like to transform is decomposed into a set of appropriate basis functions. Instead of using trigonometric 
functions as in the FT we use Dirac combs (DC) with 
different inter-spike periods as basis functions, Eq.~\eqref{eq_dirac_comb}. Let $s(t_i)$ be the normalized signal we want to transform of length $N$ and 
$t_i=i\cdot \Delta t,~i=1,\ldots,N$, where $\Delta t$ denotes the sampling time and $s(t_i) \in [0,\ 1]\ \forall\ i$. In the following we label this time series as a $(1\times N)$-dimensional 
vector $\vec{s}$. 
First, $\tilde{N}$ different DCs of length $N$ are constructed with inter-spike periods $T_\text{is} \in [1,\ldots,\tilde{N}]$ and $\tilde{N}=\ceil*{N/2}+1$. Second, in order to account 
for possible phase shifts of 
these basis functions occurring in $\vec{s}$, each of these $\tilde{N}$ different DCs also need to be shifted one step further $T_\text{is}-1$ times. This leaves us with a total number of 
$M = \sum_{i=1}^{\tilde{N}}i$ 
basis functions which we can arrange as rows of a $(M\times N)$-sized matrix $\bfX$
(Fig.~\ref{fig_tau_rr_basis_functions} illustrates the described procedure):
\begin{align}
\label{eq_basis_matrix} 
\bfX_{i,j} &= \sum_{k=0}^{N}\delta\bigl(j-1-kT_\text{is}(i)-i+T_\text{is}(i) \bigr), \quad i=1,\dots,\ceil*{N/2}+1, \quad j=1,\dots,N\\
\label{eq_T_is} T_\text{is}(i)&= n, \quad \text{such that}\quad n:~ \frac{n(n-1)}{2} + 1 \leq i <  \frac{n(n+1)}{2} + 1,\quad n \in \mathbb{N}_+.
\end{align}

@@

\noindent Eventually, an under-determined linear system
\begin{equation}
\bfX^T \boldsymbol{\beta}=\vec{s}
\label{eq_linear_system}
\end{equation}  
has to be solved for $\boldsymbol\beta$, a $(M\times 1)$-sized vector carrying the loadings we are interested in. Along a variety of algorithms which can solve this problem, we are 
particularly interested in those solutions, which promote sparsity in $\boldsymbol\beta$, since our goal is to decompose the signal $\vec{s}$ into a minimal number of basis 
functions (for an excellent overview of the topic we refer to \citet{Brunton2019}). In this paper we either use the \textit{least absolute shrinkage and selection operator} 
(LASSO) \cite{Tibshirani1996} or a \textit{sequentially thresholded least squares} (STLS) regression \cite{Brunton2016,Brunton2019} to obtain a solution $\hat{\boldsymbol\beta}$. Finally, we group loadings 
which correspond to basis functions having the same period 
$T_\text{is}$ into $\hat{\boldsymbol\beta}_f$ and obtain the 
\textit{inter-spike spectrum} by simply plotting $\hat{\boldsymbol\beta}_f$ as a function of the frequency $f=T_\text{is}^{-1}$, with $T_\text{is}=\Delta t, 2\Delta t, \dots, (\ceil*{N/2}+1) \Delta t$ 
(Fig.~\ref{fig_tau_rr_dirac_comb}C, F).

@@

When introducing a sparsity regularization we give up the ability to fully reconstruct the signal from the decomposition in favor of identifying the most important spikes, and thus the most important periodicities more clearly. This choice is parametrized by a regression regularization parameter $\alpha$. Only for $\alpha = 0$ do we obtain a solution to Eq.~(\ref{eq_linear_system}) and thus a full reconstruction of the signal. Our general approach to fixing $\alpha$ is to require that we preserve most of the original siginal in the sense of matching a given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and the reconstruction.

%
%In addition to the computational complexity of solving Eq.~\eqref{eq_linear_system}, the solving algorithm must also deal with the free choice of the regression regularization parameter $\alpha$, 
%which determines the sparsity of the regression coefficient vector $\hat{\boldsymbol\beta}$. Obviously, there will be numerous peaks in the inter-spike spectrum of a given signal, when $\alpha$ is small, 
%since $\alpha=0$ corresponds to the least squares solution and the corresponding $\hat{\boldsymbol\beta}$ is an exact solution of Eq.~(\ref{eq_linear_system}) and not sparse. On the other hand, 
%only a sparse number of non-zero loadings in $\hat{\boldsymbol\beta}$, i.e., peaks in the inter-spike spectrum, are obtained when $\alpha$ is sufficiently large. 
%The actual implementation of the regularization depends on the regression method chosen (in our case, LASSO or STLS), and consequently a given value of $\alpha$ will generally result in 
%different sparse loading vectors $\hat{\boldsymbol\beta}$. Therefore, in this paper we select $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
%given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and itself.

@@

We have successfully used the idea of transforming the $\tau$-RR for the detection of bifurcations in the Logistic map. By constructing appropriate surrogates of the inter-spike spectra, 
and thus a null model, the number of significant peaks in the inter-spike spectrum correlated well with the positive Lyapunov exponent. This measure was also able to resolve 
period-doubling bifurcations. However, we have to admit this way the detection of a bifurcation is only possible when the additional period(s) is not an integer multiple of the former period(s). 
This behavior is described in the application to the R\"ossler system, where we explicitly showed the different inter-spike power spectra for period-2, period-3, and chaotic dynamics.
Further development might potentially incorporate the mutual distance of peaks in the spectrum for a better correlation to the Lyapunov exponent. The inter-spike spectra of power grid 
frequency data illustrate that our proposed method may serve as a valuable source of information in addition to a standard Fourier analysis even for scalar time series. And last but not least, we showed that our approach 
really incorporates frequencies, which are apparent in the Fourier spectrum of the signal, by applying it to analytically derived eccentricity data, where the dominant frequencies are well 
known.

We discuss some more technical details in the following, which will also affect any application of the proposed method. 
First of all, the number of required basis functions 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ for an input signal of length $N$ is the crucial bottleneck of this approach, which is why it does not show good scaling behavior. The subsequent sparse 
regression, therefore, gets computationally intensive for $N>1,000$. Depending on the memory of the computer used, input signals $N>2,000$ usually do not work anymore. This means that signals 
often need to get downsampled as a preprocessing step (e.g., see Appendix \ref{sec_power_grid_appendix}). Second, the regularization parameter $\alpha$ for the regression is a crucial free parameter. 
As described in Sect.~\ref{sec_tau_rr_method}, our idea in this paper is to select the $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
given (Pearson) correlation coefficient $\rho_{\vec{s},\tilde{\vec{s}}}$ between the original signal $\vec{s}$ and itself. This ensures that $\alpha$ adjusts itself to the data as well as to the 
used regression method. We found that this increases the comparability of different spectra, especially when performing a running window approach in order to obtain an evolutionary spectrogram 
(Fig.~\ref{fig_laskar_spectra}). 
However, the two different sparse regression algorithms we encountered in this article (LASSO and STLS) yield different results for the same desired $\rho_{\vec{s},\tilde{\vec{s}}}$. 
Even if the spectra obtained in this way look qualitatively similar, they are not always quantitatively similar. The reason for this 
is that $\rho_{\vec{s},\tilde{\vec{s}}}$ is not a smooth function of $\alpha$ in case of STLS, due to the hard-thresholding involved \cite{Brunton2016}, which is shown in Fig.~\ref{fig_regularization}.
Third, when adopting our idea of applying the inter-spike spectrum to the $\tau$-recurrence rate of the signals state space trajectory, the embedding process induces additional free parameters. 
This is not a drawback of the proposed method, but rather a drawback of the particular application method that we have heavily used in this article and which was the main motivation for developing 
the proposed method. As a very last remark, we draw attention to the fact that sparse regression can be transformed into sparse logistic regression when the signal we would like to transform is binary.

@@

A novel type of power spectral density, the \textit{inter-spike spectrum}, has been proposed. The method decomposes any arbitrary signal into basis functions which consist of (lagged) Dirac 
combs (DC) of different inter-spike period. The loading for each period is obtained by a regularized regression, which promotes sparsity in its solution. We chose LASSO or a sequentially 
thresholded least squares regression STLS in this letter. Since there are 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ basis functions for a signal of length $N$ the regression can get computationally intensive for $N>1,000$. When plotting the computed loadings as a function 
of the period (or frequency) the inter-spike spectrum is obtained. This provides information that is complementary to the Fourier decompostion of the system.

>>>>

% Abstract (Do not insert blank lines, i.e. \\)

\abstract{In recurrence analysis, the $\tau$-recurrence rate encodes the periods of the cycles of the underlying high dimensional time series.
It thus plays a similar role to the autocorrelation for scalar time series in encoding temporal correlations.
Unlike the autocorrelation however, its Fourier decomposition does not have a natural interpretation.
Thus there is no direct analogue to the power spectrum in recurrence analysis.
In this paper we introduce a novel way to decompose the $\tau$-recurrence rate using an over-complete basis of Dirac combs together with a sparsity regularization.
We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis in the sense that it reveals the dominant periodicities of the underlying time series.
We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying system in a wide variety of examples and is robust to measurement noise.}

%
%A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
%clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
%We invent this instructive approach with the 
%idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
%along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
%spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}

@@

\citet{Zbilut2008} pointed out that $\tau$-RR could be interpreted as analogous to the auto-correlation function $C(\tau)$ for a scalar time series.
However, as there is no equivalent of the Wiener-Khinchin theorem its Fourier Transform (FT) only provides a ``generalized'' kind of power spectral density.
There are clear advantages to such a recurrence-derived power spectral density, the (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D,E):
There are no assumptions for stationarity or sampling, when constructing a RP, 
and since an RP can represent high dimensional dynamics the approch directly applies to higher dimensional systems for which the auto-correlation is a matrix and no single power spectrum captures the recurrences.\\

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Fig.~\ref{fig_tau_rr_dirac_comb}A, D), so 
in principle we would strive for a single peak in the frequency domain of this signal at a frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many frequencies excited (Fig.~\ref{fig_tau_rr_dirac_comb}B, E). This is because the Fourier components add constructively 
for every frequency $1/T_\text{is}$ and therefore $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$ \cite{Norden1998}.\footnote{This can also 
be observed for neuron spike trains, e.g. \cite{Orcioni2020,Biagetti2017}.}

@@

In this article we propose a new way of decomposing a spike-train-like signal into periodic components as 
an alternative to the RP-based method suggested in \cite{Zbilut2008} or Fourier-based spike-train power 
spectra \cite{Dummer2014}. We instead investigate decomposing the $\tau$-RR in an overcomplete basis of spike trains using a sparsity regularization. Section \ref{sec_tau_rr_method} introduces the approach in detail. Our novel \textit{inter-spike spectrum} does not show resonance 
behavior of the signal's inherent inter-spike frequencies (Fig.~\ref{fig_tau_rr_dirac_comb}C, F). Note that this approach can be used to decompose arbitrary signals, and is not specific to $\tau$-RR. The more spiky the signal is, the more useful our new decomposition can potentially be. In Section \ref{sec_tau_rr_application} we demonstrate the usage of inter-spike spectrum when transforming the $\tau$-RR of a system under study. In this case, the inter-spike spectrum 
can unravel characteristic time scales of high dimensional systems, which is not possible when using a FT. Finally, in Section \ref{sec_tau_rr_conclusion} our results are summarized.

@@

To obtain the Inter-Spike Spectrum , the signal, which we would like to analyze, in our case the $\tau$-RR, is decomposed into a set of appropriate basis functions. 
The general idea common to many methods is that the sum of these weighted functions can approximate a finite signal to a sufficient degree. The weights (in some contexts also called modes or loadings) corresponding to the individual basis functions must be determined.
A number of decomposition techniques based on different sets of basis functions exist, e.g., 
trigonometric functions (Fourier- and wavelet analysis \cite{Bracewell1986}), 
eigenvectors of the corresponding covariance matrix (Principal Component Analysis \cite{Hotelling1933} and related techniques) 
or Intrinsic Mode Functions (Empirical Mode Decomposition and Hilbert spectrum \cite{Norden1998}).
These methods typically share the property that the basis functions form a complete basis,
the set of basis functions is linearly independent, and thus the weights are uniquely determined.

Here we propose to use Dirac combs (DC) with 
different inter-spike periods as basis functions, Eq.~\eqref{eq_dirac_comb}. Let $s(t_i)$ be the min-max-normalized signal we want to transform of length $N$ and 
$t_i=i\cdot \Delta t,~i=1,\ldots,N$, where $\Delta t$ denotes the sampling time and $s(t_i) \in [0,\ 1]\ \forall\ i$. In the following we label this time series as a $(1\times N)$-dimensional 
vector $\vec{s}$. 
First, $\tilde{N}$ different DCs of length $N$ are constructed with inter-spike periods $T_\text{is} \in [1,\ldots,\tilde{N}]$ and $\tilde{N}=\ceil*{N/2}+1$. Second, in order to account 
for possible phase shifts of 
these basis functions occurring in $\vec{s}$, each of these $\tilde{N}$ different DCs also need to be shifted one step further $T_\text{is}-1$ times. This leaves us with a total number of 
$M = \sum_{i=1}^{\tilde{N}}i$ 
basis functions which we can arrange as rows of a $(M\times N)$-sized matrix $\bfX$
(Fig.~\ref{fig_tau_rr_basis_functions} illustrates the described procedure):
\begin{align}
\label{eq_basis_matrix} 
\bfX_{i,j} &= \sum_{k=0}^{N}\delta\bigl(j-1-kT_\text{is}(i)-i+T_\text{is}(i) \bigr), \quad i=1,\dots,\ceil*{N/2}+1, \quad j=1,\dots,N\\
\label{eq_T_is} T_\text{is}(i)&= n, \quad \text{such that}\quad n:~ \frac{n(n-1)}{2} + 1 \leq i <  \frac{n(n+1)}{2} + 1,\quad n \in \mathbb{N}_+.
\end{align}

@@

\noindent Eventually, an under-determined linear system
\begin{equation}
\bfX^T \boldsymbol{\beta}=\vec{s}
\label{eq_linear_system}
\end{equation}  
has to be solved for $\boldsymbol\beta$, a $(M\times 1)$-sized vector carrying the weights we are interested in. Along a variety of algorithms which can solve this problem, we are 
particularly interested in those solutions, which promote sparsity in $\boldsymbol\beta$, since our goal is to decompose the signal $\vec{s}$ into a minimal number of basis 
functions (for an excellent overview of the topic we refer to \citet{Brunton2019}). In this paper we either use the \textit{least absolute shrinkage and selection operator} 
(LASSO) \cite{Tibshirani1996} or a \textit{sequentially thresholded least squares} (STLS) regression \cite{Brunton2016,Brunton2019} to obtain a solution $\hat{\boldsymbol\beta}$. Any other 
sparse regression technique could be used. Finally, we group weights 
which correspond to basis functions having the same period 
$T_\text{is}$ into $\hat{\boldsymbol\beta}_f$ and obtain the 
\textit{inter-spike spectrum} by simply plotting $\hat{\boldsymbol\beta}_f$ as a function of the frequency $f=T_\text{is}^{-1}$, with $T_\text{is}=\Delta t, 2\Delta t, \dots, (\ceil*{N/2}+1) \Delta t$ 
(Fig.~\ref{fig_tau_rr_dirac_comb}C, F).

@@

When introducing a sparsity regularization we give up the ability to fully reconstruct the signal from the decomposition in favor of identifying the most prominent spikes, and thus the most important periodicities more clearly. This choice is parametrized by a regression regularization parameter $\alpha$. Only for $\alpha = 0$ do we obtain a solution to Eq.~(\ref{eq_linear_system}) and thus a full reconstruction of the signal. Our general approach to fixing $\alpha$ is to require that we preserve most of the original signal in the sense of matching a given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and the reconstruction. The chosen regression method (in our case, LASSO or STLS) will determine how sensitive the obtained inter-spike spectrum is with respect to the regularization given by a 
selected $\rho$ (Fig.~\ref{fig_regularization}). Regarding this, in most of our applications LASSO yields more robust results, while STLS is usually faster to compute. 
We will discuss this in more detail in Section 4. Section \ref{sec_tau_rr_discussion}.

%
%In addition to the computational complexity of solving Eq.~\eqref{eq_linear_system}, the solving algorithm must also deal with the free choice of the regression regularization parameter $\alpha$, 
%which determines the sparsity of the regression coefficient vector $\hat{\boldsymbol\beta}$. Obviously, there will be numerous peaks in the inter-spike spectrum of a given signal, when $\alpha$ is small, 
%since $\alpha=0$ corresponds to the least squares solution and the corresponding $\hat{\boldsymbol\beta}$ is an exact solution of Eq.~(\ref{eq_linear_system}) and not sparse. On the other hand, 
%only a sparse number of non-zero weights in $\hat{\boldsymbol\beta}$, i.e., peaks in the inter-spike spectrum, are obtained when $\alpha$ is sufficiently large. 
%The actual implementation of the regularization depends on the regression method chosen (in our case, LASSO or STLS), and consequently a given value of $\alpha$ will generally result in 
%different sparse loading vectors $\hat{\boldsymbol\beta}$. Therefore, in this paper we select $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
%given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and itself.

@@

We have successfully used the idea of transforming the $\tau$-RR for the detection of bifurcations in the Logistic map. By constructing appropriate surrogates of the inter-spike spectra, 
and thus a null model, the number of significant peaks in the inter-spike spectrum correlated well with the positive Lyapunov exponent. This measure was also able to resolve 
period-doubling bifurcations. However, we have to admit this way the detection of a bifurcation is only possible when the additional period(s) is not an integer multiple of the former period(s). 
This behavior is described in the application to the R\"ossler system, where we explicitly showed the different inter-spike spectra for period-2, period-3, and chaotic dynamics.
Further development might potentially incorporate the mutual distance of peaks in the spectrum for a better correlation to the Lyapunov exponent. The inter-spike spectra of power grid 
frequency data illustrate that our proposed method may serve as a valuable source of information in addition to a standard Fourier analysis even for scalar time series. And last but not least, we showed that our approach 
really incorporates frequencies, which are apparent in the Fourier spectrum of the signal, by applying it to analytically derived eccentricity data, where the dominant frequencies are well 
known.

We discuss some more technical details in the following, which will also affect any application of the proposed method. 
First of all, the number of required basis functions 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ for an input signal of length $N$ is the crucial bottleneck of this approach, which is why it does not show good scaling behavior. The subsequent sparse 
regression, therefore, gets computationally intensive for $N>1,000$. Depending on the memory of the computer used, input signals $N>2,000$ usually do not work anymore. This means that signals 
often need to get downsampled as a preprocessing step (e.g., see Appendix \ref{sec_power_grid_appendix}). Second, the regularization parameter $\alpha$ for the regression is a crucial free parameter. 
As described in Sect.~\ref{sec_tau_rr_method}, our idea in this paper is to select the $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
given (Pearson) correlation coefficient $\rho_{\vec{s},\tilde{\vec{s}}}$ between the original signal $\vec{s}$ and itself. This ensures that $\alpha$ adjusts itself to the data as well as to the 
used regression method. We found that this increases the comparability of different spectra, especially when performing a running window approach in order to obtain an evolutionary spectrogram 
(Fig.~\ref{fig_laskar_spectra}). 
However, the two different sparse regression algorithms we encountered in this article (LASSO and STLS) yield different results for the same desired $\rho_{\vec{s},\tilde{\vec{s}}}$. 
Even if the spectra obtained in this way look qualitatively similar, they are not always quantitatively similar. The reason for this 
is that $\rho_{\vec{s},\tilde{\vec{s}}}$ is not a smooth function of $\alpha$ in case of STLS, due to the hard-thresholding involved \cite{Brunton2016}, which is shown in Fig.~\ref{fig_regularization}.
Third, when adopting our idea of applying the inter-spike spectrum to the $\tau$-recurrence rate of the signals state space trajectory, the embedding process induces additional free parameters.  
This is not a drawback of the proposed decomposition method, but rather a drawback of applying this technique to the $\tau$-recurrence rate of the system, which was the main motivation for 
developing the proposed method. As a very last remark, we draw attention to the fact that sparse regression can be transformed into sparse logistic regression when the signal we would like to transform is binary.

@@

A novel decomposition technique is proposed that yields the \textit{inter-spike spectrum}. The method decomposes any arbitrary signal into basis functions which consist of (lagged) Dirac 
combs (DC) of different inter-spike period. The loading for each period is obtained by a regularized regression, which promotes sparsity in its solution. We chose LASSO or a sequentially 
thresholded least squares regression STLS in this letter. Since there are 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ basis functions for a signal of length $N$ the regression can get computationally intensive for $N>1,000$. When plotting the computed weights as a function 
of the period (or frequency) the inter-spike spectrum is obtained. This provides information that is complementary to the Fourier decompostion of the signal.
